---
title: "STA440 Final Project"
author: "Matthew O'Donnell"
date: December 2024
format: pdf
execute:
  echo: false
  warning: false
  message: false
---

```{r}
source("Genderedness.R", local = knitr::knit_global())
source("Weighted_Regression.R", local = knitr::knit_global())
```

# 1. Introduction 

# 2. Methodology

## 2.1 Gender-Specific Name Usage Over Time

For each character in the dataset, a gender score is provided, which is computed as follows: 

$$g_i = \frac{N_{\text{male}} - N_{\text{female}}}{N_{\text{male}} + N_{\text{female}}}$$
This gender score for a character $g_i$ can range from -1 (completely feminine) to 0 (gender-neutral) to 1 (completely masculine). The data contains a parts-per-million metric for each character during each of six birthing cohorts/generations. A average of gender specificity of a cohort (absolute value of gender score), weighted by parts-per-million in that cohort, could be used to see changes in gender-specificity over time. 

```{r}
plot
```

## 2.2 Subjective Name Characteristics By Gender

# 3. Results 

## 3.1 Anova Results

#### Coefficients Table

| Predictor                | Slope (Estimate) | p-value     |
|--------------------------|------------------|-------------|
| Intercept               | 0.593         | < 0.001     |
| Time Period : 1960-1969 | -0.003        | 0.683       |  
| Time Period :1970-1979 | 0.003         | 0.642       | 
| Time Period : 1980-1989 | 0.011         | 0.144       |  
| Time Period : 1990-1999 | -0.004        | 0.597       |  
| Time Period : 2000-2008 | -0.042        | < 0.001    |

### ANOVA Results
| Independent Variable | F-statistic | P-value |
|---|---|---|
| Time Period | 10.792 | < 0.001

## 3.2 WLS Output

### Valence
| Decade       | Estimated Coefficient ($\hat{\beta}_1$) | P-value         |
|--------------|-----------------|-----------------|
| 1930–1959    | -0.053          | < 0.001         |
| 1960–1969    | -0.026          | 0.0436          |
| 1970–1979    | -0.009          | 0.494           |
| 1980–1989    |  0.001          | 0.907           |
| 1990–1999    |  0.015          | 0.251           |
| 2000–2008    |  0.027          | 0.0394          |

### Warmth
| Decade       | Estimated Coefficient ($\hat{\beta}_1$) | P-value         |
|--------------|-----------------|-----------------|
| 1930–1959    | -0.054          | < 0.001         |
| 1960–1969    | -0.038          | 0.0007          |
| 1970–1979    | -0.032          | 0.0035          |
| 1980–1989    | -0.026          | 0.0132          |
| 1990–1999    | -0.022          | 0.0359          |
| 2000–2008    | -0.037          | 0.0010          |

### Competence
| Decade       | Estimated Coefficient ($\hat{\beta}_1$) | P-value         |
|--------------|-----------------|-----------------|
| 1930–1959    |  0.088          | < 0.001         |
| 1960–1969    |  0.134          | < 0.001         |
| 1970–1979    |  0.172          | < 0.001         |
| 1980–1989    |  0.202          | < 0.001         |
| 1990–1999    |  0.204          | < 0.001         |
| 2000–2008    |  0.206          | < 0.001         |




# 4. Discussion

